---
title: "Visualizing Target Pipelines with Mermaid"
author: "Your Package Name"
description: "How to use knit_targets_mermaid() to create interactive, zoomable, and tooltip-rich pipeline diagrams in R Markdown and Quarto."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Target Pipelines with Mermaid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The `knit_targets_mermaid()` function provides a powerful way to visualize your [`targets`](https://docs.ropensci.org/targets/) pipeline as an interactive [Mermaid](https://mermaid-js.github.io/mermaid/#/) diagram. This vignette demonstrates how to use it in R Markdown or Quarto, including zooming, panning, and rich tooltips with markdown support.

# Basic Usage

```{r, eval=FALSE}
library(yourpackagename)
# Basic pipeline
knit_targets_mermaid(targets_only = TRUE)
```

This will render a vertical (top-to-bottom) Mermaid diagram of your pipeline, with zoom and pan enabled.

# Customizing the Diagram

You can pass arguments to both `targets::tar_mermaid()` and `targets::tar_manifest()` via `...`. For example, to use a custom pipeline script:

```{r, eval=FALSE}
knit_targets_mermaid(script = "my_targets.R", targets_only = TRUE)
```

# Tooltips with Markdown

If you add a `description` to your targets, it will appear as a tooltip when you hover over the node. Descriptions are parsed as markdown (if the `markdown` package is installed) and rendered as HTML, so you can use formatting, links, and more:

```r
# In your _targets.R
list(
  tar_target(x, 1 + 1, description = "**Bold** _italic_ [link](https://r-project.org)"),
  tar_target(y, x * 2, description = "`y` is double x.\n- Bullet 1\n- Bullet 2")
)
```

# Example Output

Below is an example of a rendered diagram with tooltips:

```{r, eval=FALSE}
knit_targets_mermaid()
```

Hover over the nodes to see the markdown-formatted tooltips. You can zoom and pan the diagram with your mouse.

# Styling and Customization

The function injects custom CSS and JS for zooming, panning, and tooltips. You can further customize the appearance by editing `inst/assets/mermaid-zoom.css` and `inst/assets/mermaid-zoom.js` in your package.

# Troubleshooting

- If tooltips do not render as HTML, ensure the `markdown` package is installed.
- If you use Quarto, ensure external JS/CSS is allowed in your document settings.

# Conclusion

`knit_targets_mermaid()` makes it easy to create beautiful, interactive, and informative pipeline diagrams for your R projects. Use markdown in your target descriptions for rich tooltips, and enjoy seamless navigation with zoom and pan.
