---
title: "Visualizing Target Pipelines with Mermaid"
author: "Your Package Name"
description: "How to use knit_targets_mermaid() to create interactive, zoomable, and tooltip-rich pipeline diagrams in R Markdown and Quarto."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Target Pipelines with Mermaid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The `knit_targets_mermaid()` function provides a powerful way to visualize your [`targets`](https://docs.ropensci.org/targets/) pipeline as an interactive [Mermaid](https://mermaid-js.github.io/mermaid/#/) diagram. This vignette demonstrates how to use it in R Markdown or Quarto, including zooming, panning, and rich tooltips with markdown support.

## Basic Usage

Create a chunk in your R Markdown or Quarto document and call `knit_targets_mermaid()` with chunk options `results = "asis"` to render the diagram directly in the document:

```{r, eval=FALSE}
library(myworkspace)
# Basic pipeline
knit_targets_mermaid(targets_only = TRUE)

```

Comparing with the default `targets::tar_mermaid()`, the following items are added:

* Change from horizontal (left-to-right) to vertical (top-to-bottom) layout.
* Add config to reduce the space between nodes and edges.
* Add tooltips that support markdown formatting.

## Customizing the Diagram

You can pass arguments to both `targets::tar_mermaid()` and `targets::tar_manifest()` via `...`. For example, to use a custom pipeline script:

```{r, eval=FALSE}
knit_targets_mermaid(script = "my_targets.R", targets_only = TRUE)
```

## Tooltips with Markdown

If you add a `description` to your targets, it will appear as a tooltip when you hover over the node. Descriptions are parsed as markdown (if the `markdown` package is installed) and rendered as HTML, so you can use formatting, links, and more:

```r
# In your _targets.R
list(
  tar_target(x, 1 + 1, description = "**Bold** _italic_ [link](https://r-project.org)"),
  tar_target(y, x * 2, description = "`y` is double x.\n- Bullet 1\n- Bullet 2")
)
```

## Example Output

Below is an example of a rendered diagram with tooltips:

```{r, eval=TRUE, out.width='100%', results='asis'}
script_file <- system.file("examples", "_targets.R", package = "myworkspace")
myworkspace::knit_targets_mermaid(script = script_file, targets_only = TRUE)
```

Hover over the nodes to see the markdown-formatted tooltips. You can zoom and pan the diagram with your mouse.

